---
title: "BRC Model 0007"
author: "Leyla Azarang and Renee Menezes" 
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    toc_collapsed: true
  toc_depth: 3  
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```


```{r, include=FALSE}
if (!require('pacman')) install.packages('pacman'); library(pacman)
```



```{r, include=FALSE}
p_load("survival", "survminer", "tidyverse", "JMbayes2")
```



```{r}
source("./scripts/mysplit.R") 
source("./scripts/functions_data_centre.R")
```

## Read in data

First read in information about centres available for this model.

```{r}
xi <- 10
alldata <- readxl::read_excel("./data/overview1.xlsx")
myfile <- alldata$FileName[xi]
mymodel <- strsplit(myfile, split = " ")[[1]][ 1 ]
all.sheets <- as.character(alldata[xi, -1])
names(all.sheets) <- colnames(alldata)[ -1 ] 
all.sheets <- all.sheets[ !is.na(all.sheets) ]
list.centers <- names(all.sheets)
```

The data here is for model `r mymodel`. It was analysed by the following centers:

```{r}
knitr::kable(list.centers, col.names = "Centers")
```


```{r echo = FALSE, eval = FALSE}
mymodel # name of the model
#all.sheets # names of the sheets to be used
list.centers # names of the centers available
```

### Analysis per center

Reading the data per center.

```{r}
n_centers <- length(list.centers)

data_centers <- lapply(1:n_centers, read_sheet, myfile = myfile, 
                       all.sheets = all.sheets)
names(data_centers) <- list.centers
``` 

Extract treatment, start and end dates from Label column. 
For this batch of analyses the event variable is not consistently recorded across all data files.

The data is currently in the wide format, and for the analyses we need it in the long format (one observation per row, for one individual at a time).



```{r}
all_data_long <- lapply(1:n_centers, get_data_long, mydatac = data_centers)
names(all_data_long) <- list.centers
```

We now need to leave out measurements before the start of the treatment, and after ending it. 

```{r}
data_long_treat <- lapply(1:n_centers, get_vol_treat, alldata = all_data_long)
names(data_long_treat) <- names(all_data_long)
```



## Exploratory data analysis

The following plots depict the tumor volume growth per each center, for each model separately. 

```{r results = 'hide'}
vp <- lapply(1:n_centers, function(xc){ 
p <-  ggplot(data = data_long_treat[[ xc ]], aes(x = days, y = volume, 
                                             group = `GenealogyID`, color =Treatment)) +
    geom_line()+
    facet_wrap( vars( `GenealogyID`)  )+
    ylab("tumor volume")+
    theme(legend.position = "top") +
    ggtitle(paste(names(data_long_treat)[xc]) )
return(p)
}
)
```

```{r fig.width = 10, fig.height = 10}
vp
```



Now group all curves per treatment.


```{r results = 'hide'}

vp <- lapply(1:n_centers, function(xc){ 
p <-  ggplot(data = data_long_treat[[ xc ]], aes(x = days, y = volume, 
                                             group = GenealogyID, color = Treatment)) +
    geom_line()+
    facet_wrap( vars( `Treatment`)  )+
    ylab("tumor volume")+
    theme(legend.position = "top") +
    ggtitle(paste(names(data_long_treat)[xc]) )
return(p)
}
)
```

```{r fig.width = 10, fig.height = 6}
vp
```


## The  impact of center on tumor growth 

In this section we investigate the effect of center on tumor growth  in each treatment level. For this, we put all data together into a single data frame.


```{r}
all_data <- NULL
for(xc in 1:n_centers)
{
  datac <- data_long_treat[[ xc ]]
  datac$center <- rep(list.centers[ xc ], nrow(datac))
  all_data <- rbind(all_data, datac)
}
all_data <- all_data[!is.na(all_data$volume), ]
```








## The effect of treatment per center


```{r}
cont <- nlmeControl( maxIter = 100000, msMaxIter = 1000, tolerance = 1e-6, niterEM = 10000, msMaxEval = 10000)

all_fits <- lapply(1:n_centers, function(xc){ 
      lme_fit <- lme(volume ~   days +  days:Treatment, data = data_long_treat[[xc]], 
                     random = ~ days | GenealogyID, control = cont)
      mysumm <- summary(lme_fit)
      mfitted <- fitted(lme_fit, level =0)
      res_fitted <- tibble( GenealogyID = names(mfitted), 
                           Treatment = data_long_treat[[ xc ]]$Treatment, 
                            days = data_long_treat[[ xc ]]$days, fit_fix = mfitted)
      
      return(list( lme_fit= lme_fit , fit_sum = mysumm , fit_tibble = res_fitted  ) )
}
)

list_lme_fit <- lapply(1:n_centers , function(xc){return(all_fits[[xc]][[1]])})
names(list_lme_fit) <- names(data_long_treat)

list_lme_summ <- lapply(1:n_centers , function(xc){return(all_fits[[xc]][[2]])})
names(list_lme_summ) <- names(data_long_treat)

list_lme_fitted <- lapply(1:n_centers , function(xc){return(all_fits[[xc]][[3]])})
names(list_lme_fitted) <- names(data_long_treat)
```








The following plots depict the tumor volume increase in each  center for both  "PBS" and "TDM" treatments.

```{r}
vp <- lapply(1:n_centers, function(xc){ 
    
p <- ggplot(data = list_lme_fitted[[xc]], aes(x = days, y = fit_fix))+
  geom_line(aes(color=Treatment))+
  ylab("tumor volume")+
  ggtitle(paste(names(data_long_treat)[xc]) )
return(p)
}
)
```

```{r}
vp
```


## Survival analysis

Below we make a list of survival data corresponding to different center. 

```{r}
surv_data_centers <- lapply(1:n_centers , get_surv_data , mydatac = data_centers)
names(surv_data_centers) <- names(data_centers)
```



```{r}
tab_event_centers <- lapply(1:n_centers, function(xc){
  tab_event = table(surv_data_centers[[xc]]$`event/status/sensor`)
 return(tab_event)
}
) 
  
names(tab_event_centers) <- names(data_centers)
```




```{r}
tab_event_centers
```



```{r}
surv_object_centers <- lapply(1:n_centers , get_surv_object ,surv_data_centers)
names(surv_object_centers) <- names(data_centers)
```


```{r}
surv_fit_centers <- lapply(1:n_centers , get_surv_fit, surv_object_centers , surv_data_centers)
names(surv_fit_centers) <- names(data_centers)
```


From the following tables, for each center, we can see the number of events in each treatment group. 

```{r}
surv_fit_centers
```


### The effect of treatment on survival time 

The following plots depict the survival curves in each center. 


```{r}
lapply( 1:n_centers, function(xc){
  plot(surv_fit_centers[[xc]], 
     ylab = "Survival Proportion", 
     xlab = "Time",
     col = c(1,2),
     main = paste("KM curve",names(data_centers)[xc])) 
    }
)
```



```{r}
cox_centers <- lapply(1:n_centers, function(xc){ 
    cox <- coxph(surv_object_centers[[xc]] ~ 1 , data = surv_data_centers[[xc]])
    return(cox)
}
)
names(cox_centers)<- names(data_centers)
```


## Joint modelling

In this section, we use both survival and tumor growth altogether  to analyse the data. In this part, for each center,  we investigate the effect of the treatment on tumor growth  by using the the association of tumor growth with survival time. 




```{r}
jointFit_centers <- lapply(1:n_centers, function(xc){ 
    jointFit <- jm(cox_centers[[xc]], list_lme_fit[[xc]], time_var = "days", data_Surv = surv_data_centers[[xc]] ,  n_iter = 85000L, n_burnin = 3500L)
    sum <- summary(jointFit)
    return(sum)
}
)
names(jointFit_centers) <- names(data_centers)
```


```{r}
jointFit_centers
```


```{r}
Name_Treat <- levels(factor(unique(data_long_treat[[5]]$Treatment)))
list_data_JMfit <- lapply(1:n_centers , function(xc){
  fit <- jointFit_centers[[xc]]$Outcome1
 a <- fit[1,"Mean"]
 b_Cont <- fit["days","Mean"]
 b_Treat <- b_Cont + fit[3,"Mean"] 

 days <- 0:max(0:max(data_long_treat[[xc]]$days))
 pred_Cont <- a + b_Cont *days
 pred_Treat <- a + b_Treat *days
 Group <- rep(names(data_long_treat)[xc], length(days))
 data_JMfit <- tibble(day = days, pred_Cont= pred_Cont,       pred_Treat= pred_Treat, Group= Group)
 
 data_JMfit <- data_JMfit %>%
  gather(key= "Treatment", value="Pred", -day,  -Group ) %>%
  mutate(Pred = as.numeric(Pred))%>%
  mutate(Treatment =factor( ifelse(Treatment =="pred_Cont", Name_Treat[1], Name_Treat[2])))
 return(data_JMfit)
}
)

names(list_data_JMfit) <- names(data_centers)
```



```{r}
# plot joint models
list_data_JMfit <- lapply(1:n_centers , function(xc){
  fit <- jointFit_centers[[xc]]$Outcome1
 a <- fit[1,"Mean"]
 b_Cont <- fit["days","Mean"]
 b_Treat <- b_Cont + fit[3,"Mean"] 

 days <- 0:max(0:max(data_long_treat[[xc]]$days))
 pred_Cont <- a + b_Cont *days
 pred_Treat <- a + b_Treat *days
 Group <- rep(names(data_long_treat)[xc], length(days))
 data_JMfit <- tibble(day = days, pred_Cont= pred_Cont,       pred_Treat= pred_Treat, Group= Group)
 
 data_JMfit <- data_JMfit %>%
  gather(key= "Treatment", value="Pred", -day,  -Group ) %>%
  mutate(Pred = as.numeric(Pred))%>%
  mutate(Treatment =factor( ifelse(Treatment =="pred_Cont", "PBS", "TDM")))
 return(data_JMfit)
}
)

names(list_data_JMfit) <- names(data_centers)
pvals <- round(sapply(1:n_centers, function(xc)jointFit_centers[[xc]]$Outcome1[["days:TreatmentTDM", "P"]]),3)
```


```{r}
JMp <- lapply(1:n_centers, function(xc){ 
    
p <- ggplot(data =list_data_JMfit[[xc]], aes(x = day, y = Pred, color = Treatment))+
   geom_line()+
   geom_hline( yintercept = 1500, color="#331112", linetype="longdash")+
   ylab("tumor volume")+
   scale_x_continuous( breaks=seq(0,45,5)) +
  scale_y_continuous(lim=c(0,2000),breaks=seq(0,2000,500))+
 ggtitle(paste(names(list_data_JMfit)[xc]) )+
     annotate("text", x = 20, y = 100, label = paste("p-value:", pvals[[xc]]) ) 
return(p)
}
)
```



```{r}
JMp
```


## Conclusion


By taking  both survival information and tumor volume data into consideration,  we observe a significant treatment effect over time in all  centers except "KULEUVEN".  
Compared to vehicle group, in treatment group  the tumor growth is  significantly and substantially slower, but in "KULEUVEN" the tumor growth  in treatment group is slightly smaller  than that in vehicle group, and even this small difference is not  significant. 
