---
title: " Comparision CUP Model 0006 and orthotopic 0006"
author: "Leyla Azarang and Renee Menezes" 
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    toc_collapsed: true
  toc_depth: 3  
---







```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```




```{r, include=FALSE}
if (!require('pacman')) install.packages('pacman'); library(pacman)
```



```{r, include=FALSE}
p_load("survival", "survminer", "tidyverse", "JMbayes2")
```



```{r}
source("./scripts/mysplit.R") 
source("./scripts/functions_data_centre_ortho.R")
```

## Load data


```{r}
load("save/all_data_s_0006")
all_data_s <- all_data %>%
    mutate(type = "standard")
load("save/all_data_o_0006")
all_data_o <- all_data %>%
    mutate(type = "orthotopic")
```



```{r}
all_data_s_o <- bind_rows (all_data_s, all_data_o) %>%
    mutate(type = factor(type)) %>%
    mutate(Treatment = factor(ifelse(Treatment == "PBS" | Treatment == "Physio" , "PBS", "Cetuximab"), levels = c("PBS","Cetuximab")))

```

# The effect of treatment on tumor growth per study

Regardless of the center we want to see if there is a difference between the effect of treatment on tumor volume between standard and orthotopic experiments. 


```{r}
cont <- nlmeControl( maxIter = 100000, msMaxIter = 1000, tolerance = 1e-6, niterEM = 10000, msMaxEval = 10000)

lme_fit <- lme(volume ~   days  + type +  days:Treatment  + days:type + days:type: Treatment, data = all_data_s_o, random = ~ days | GenealogyID, control = cont)

summary(lme_fit)
```

```{r}
fitted <- fitted(lme_fit, level = 0)



fitted <- tibble( MouseID = names(fitted),   type = all_data_s_o$type, 
                 Treatment = all_data_s_o$Treatment, days = all_data_s_o$days, fit_fix = fitted)
```



The following plot depicts the tumor volume increase in different studies for both  treatment groups. 

```{r}
ggplot(fitted, aes(x = days, y = fit_fix, color = type))+
  geom_line(aes(linetype=Treatment))+
  ylab("tumor volume")+
     annotate("text", x = 25, y = 500, label = paste("p-value:", 0.0000))
```

# Conclusion

- Tumor growth in standard study is significantly faster than orthotopic study.
- Unlike in standard study, in the orthotopic study the effect of treatment is not significant.  
- In standard study the treatment effect is significantly bigger than orthotopic study.

