---
title: " Comparision BRC Model 0007 and humanized 0007"
author: "Leyla Azarang and Renee Menezes" 
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    toc_collapsed: true
  toc_depth: 3  
---







```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```



```{r, include=FALSE}
if (!require('pacman')) install.packages('pacman'); library(pacman)
```



```{r, include=FALSE}
p_load("survival", "survminer", "tidyverse", "JMbayes2")
```



```{r}
source("./scripts/mysplit.R") 
source("./scripts/functions_data_centre_ortho.R")
```

# Load data


```{r}
load("save/all_data_s_0007")
all_data_s <- all_data %>%
    mutate(type = "standard") %>%
    filter (center != "KULEUVEN")

load("save/all_data_h_0007")
all_data_h <- all_data %>%
    mutate(type = "humanized")

load("save/surv_data_s_0007")
surv_data_s <- surv_data_centers 


load("save/surv_data_h_0007")
surv_data_h <- surv_data_centers
```



```{r}
all_data_s_h <- bind_rows (all_data_s, all_data_h) %>%
    mutate(type = factor(type)) %>%
    mutate(Treatment = factor(ifelse(Treatment == "PBS", "PBS", "TDM")) )
```


# The effect of treatment on tumor growth per study

Regardless of the center we want to see if there is a difference between the effect of treatment on tumor volume between standard and humanized experiments. 

```{r}
cont <- nlmeControl( maxIter = 100000, msMaxIter = 1000, tolerance = 1e-6, niterEM = 10000, msMaxEval = 10000)

lme_fit <- lme(volume ~   days  + type +  days:Treatment  + days:type   , 
               data = all_data_s_h, random = ~ days | GenealogyID, control = cont)

summary(lme_fit)
```


```{r}
cont <- nlmeControl( maxIter = 100000, msMaxIter = 1000, tolerance = 1e-6, niterEM = 10000, msMaxEval = 10000)

lme_fit <- lme(volume ~   days  + type +  days:Treatment  + days:type + days:type: Treatment  , 
               data = all_data_s_h, random = ~ days | GenealogyID, control = cont)

summary(lme_fit)
```


```{r}
fitted <- fitted(lme_fit, level = 0)



fitted <- tibble( MouseID = names(fitted),   type = all_data_s_h$type, 
                 Treatment = all_data_s_h$Treatment, days = all_data_s_h$days, fit_fix = fitted)
```



The following plot depicts the tumor volume increase in different studies for both  treatment groups. 

```{r}
ggplot(fitted, aes(x = days, y = fit_fix, color = type))+
  geom_line(aes(linetype=Treatment))+
  ylab("tumor volume") +
     annotate("text", x = 20, y = 500, label = paste("p-value:", 0.03) )

```

# survival analysis

```{r}
surv_data_s <-lapply( 1:length(surv_data_s), function(i){surv_data_s[[i]] %>% mutate(center =names(surv_data_s[i]))} )
    
surv_data_s <- do.call(rbind, surv_data_s)  %>%
    mutate(type = "standard") %>%
    mutate(event = `event/status/sensor`) %>%
    filter (center != "KULEUVEN")

surv_data_h <-lapply( 1:length(surv_data_h), function(i){surv_data_h[[i]] %>% mutate(center =names(surv_data_h[i]))} )
    
surv_data_h <- do.call(rbind, surv_data_h)  %>%
    mutate(type = "humanized") %>%
    mutate(event = `event/status`)
```


```{r}
surv_data_s_h <- bind_rows (surv_data_s, surv_data_h) %>%
    mutate(type = factor(type)) %>%
    mutate(Treatment = factor(ifelse(Treatment == "PBS", "PBS", "TDM")) ) %>%
    select(-`event/status`, -`event/status/sensor`)
```


```{r}
surv_object <- Surv(surv_data_s_h$`follow up time`, surv_data_s_h$event!=0)
```


```{r}
fit_KM <- survfit(surv_object ~ Treatment, data = surv_data_s_h)
ggsurv<- ggsurvplot(fit_KM, data = surv_data_s_h, pval = TRUE, conf.int = TRUE,  title= "KM curves")
ggsurv
```


```{r}
fit_KM <- survfit(surv_object ~ type, data = surv_data_s_h)
ggsurv<- ggsurvplot(fit_KM, data = surv_data_s_h, pval = TRUE, conf.int = TRUE,  title= "KM curves")
ggsurv
```




```{r}
cox_model <- coxph(surv_object ~ Treatment + type  , data = surv_data_s_h)
 summary(cox_model) 
```


Taking only survival information in to consideration, there is no significant difference between the survival of mice in standard and humanized studies. 



# Joint model

```{r}
set.seed(231)
jointFit <- jm(cox_model, lme_fit, time_var = "days", n_iter = 85000L, n_burnin = 3500L)
sum <- summary(jointFit)
sum
```



```{r}
# plot joint models
#Name_Treat <- levels(unique(all_data_s_h$Treatment))

fit <- sum$Outcome1
b0 <- fit[1,"Mean"]
b1 <- fit["days","Mean"]
b2type <- fit["typestandard","Mean"] 
b3treat_t <-  fit["days:TreatmentTDM","Mean"] 
b4type_t <-  fit["days:typestandard","Mean"]
b5type_t_treat <-  fit[6,"Mean"]


days <- 0:max(0:max(all_data_s_h$days))


pred_veh_hum <- b0 + b1 * days
pred_veh_stan <- b0+ b2type + (b1 + b4type_t) * days
pred_treat_hum <- b0 + (b1 + b3treat_t)* days
pred_treat_stan <- b0 + b2type + (b1 + b3treat_t + b4type_t + b5type_t_treat ) * days


data_JMfit <- tibble(day = days, pred_veh_hum = pred_veh_hum, pred_veh_stan= pred_veh_stan, pred_treat_hum = pred_treat_hum, pred_treat_stan = pred_treat_stan)
 


data_JMfit <- data_JMfit %>%
  gather(key= "treat_type", value="Pred", -day ) %>%
  mutate(Pred = as.numeric(Pred))%>%
  mutate( treatment =factor (  ifelse( treat_type =="pred_veh_hum"| treat_type =="pred_veh_stan", "PBS", "TDM"  )) ) %>%
    mutate( type =factor ( ifelse( treat_type =="pred_veh_hum"| treat_type =="pred_treat_hum", "humanized", "standard" )) )
```



```{r}
p <- ggplot(data=data_JMfit, aes(x = day, y = Pred, color = type))+
   geom_line(aes(linetype= treatment))+
   geom_hline( yintercept = 1500, color="#331112", linetype="longdash")+
   ylab("tumor volume")+
   scale_x_continuous( breaks=seq(0,45,5)) +
  scale_y_continuous(lim=c(0,2000),breaks=seq(0,2000,500))+
    theme(legend.position = "top")+
    annotate("text", x = 5, y = 1250, label = "p-value = 0.02") 
    
p 
```

# Conclusion

- Tumor growth in standard study is significantly faster than humanized study. 
- In both standard and humanized, the treatment significantly decreases tumor growth.
- However, in standard study the treatment effect is significantly bigger than humanized study. 

